### Practical Data Analysis
#### Assignement
by Jeff Leek, PhD, Roger D. Peng, PhD, Brian Caffo, PhD
Coursera June 21014 session 
  
```{r setup, include=FALSE, echo=FALSE}
# set global options
```

```{r init,echo=FALSE}
rm(list=ls())

library(knitr)
library(markdown)

library(lattice)
library(nlme)
library(ggplot2)
library(xtable)
set.seed(590607)

dt = Sys.time()
date <- format(dt,"%d-%b-%Y")
time <- format(dt,"%H:%M:%S")
rand <- rnorm(1)

#sessionInfo()
#packageVersion("snow")
Rversion <- version$version.string
```

<style>
body {
    font-size: 1.0em;
    font-family: "Arial, Helvetica, sans-serif"";
}
</style>

JMM, Document generated on `r date` at `r time`.

#### Introduction

#### Methods

This analysis has been performed using R software package for statistical analysis and the following packages.
The version of R used was `r Rversion`.

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har (section on the Weight Lifting Exercise Dataset).

Datasets:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


```{r loadData, echo=FALSE}

# b. load data

# b.1 baseDir will be prefixing all data accesses
baseDir <- ".."

# b.2 create data sub-directory if necessary
dataDir <- file.path(baseDir, "data")
if(!file.exists(dataDir)) { dir.create(dataDir) }

# b.3 download original data if necessary (skip if exists already as it takes time)
# TODO: function
fileName <- "train"
fileExt <- "csv"
fileUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
filePath <- file.path(dataDir, paste(fileName, fileExt, sep="."))
dateFilePath <- file.path(dataDir,paste(fileName,"_date_time_downloaded.txt",sep=""))
if (!file.exists(filePath)) {
  download.file (fileUrl, filePath, method="curl")
  DTDownloaded <- format(Sys.time(), "%Y-%b-%d %H:%M:%S")
  cat (DTDownloaded, file=dateFilePath)
} else {
  DTDownloaded <- scan(file=dateFilePath, what="character", sep="\n")
}
cat ("The training dataset is located at", filePath, "and was downloaded on downloaded on", DTDownloaded)

fileName <- "test"
fileExt <- "csv"
fileUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
filePath <- file.path(dataDir, paste(fileName, fileExt, sep="."))
dateFilePath <- file.path(dataDir,paste(fileName,"_date_time_downloaded.txt",sep=""))
if (!file.exists(filePath)) {
  download.file (fileUrl, filePath, method="curl")
  DTDownloaded <- format(Sys.time(), "%Y-%b-%d %H:%M:%S")
  cat (DTDownloaded, file=dateFilePath)
} else {
  DTDownloaded <- scan(file=dateFilePath, what="character", sep="\n")
}
cat ("The test dataset is located at", filePath, "and was downloaded on downloaded on", DTDownloaded)

list.files(dataDir)

# b.4 read dataset and load them in R
trainSetPath <- file.path (dataDir, "train.csv")
testSetPath <-  file.path (dataDir, "test.csv")
trainSet <- read.csv(trainSetPath, header = TRUE) 
testSet  <- read.csv(testSetPath, header = TRUE)
dim(trainSet)
dim(testSet)
str(trainSet)
str(testSet)
summary(trainSet)
summary(testSet)

#file.show("")
#head()

```

#### Results

#### Discussion

The test dataset provided was very limited in size (`r nrow(testSet)` rows) compared to the training set (`r nrow(trainSet)` rows), the test sample representing `r nrow(testSet)/((nrow(trainSet)+nrow(testSet)))`% of the total number of samples.

#### Conclusion

#### References

1. R Core Team. R: A language and environment for statistical computing. URL: http://www.R-project.org. R Foundation for Statistical Computing, 2013.

2. Seber, George AF, and Alan J. Lee. Linear regression analysis. Vol. 936. Wiley, 2012.

3. R Markdown Page. URL: http://www.rstudio.com/ide/docs/authoring/using_markdown. 

4. Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6.

5. http://groupware.les.inf.puc-rio.br/har#sbia_paper_section#ixzz33n5Wpc66
